Erik's thoughts

### 2014-11-17 WaingartenThoughts.tex
fixed typo
added more stuff

Suppose there was a function $f$ of one variable such that $F(x,y) = \dfrac{f(y)}{x}$ had periodicity 2, then $a_0 = a_2 = x$, $a_1 = y$ and
\[ a_2 = x = \dfrac{f(x)}{y} \]
And so $f(x) = xy$, for every $y$, so $f$ is not a function of one variable.

@@

Examples of functions and their periods:
\begin{itemize}
\item $F(x, y) = \dfrac{1}{x}$ with period $4$ as shown above.
\item $F(x,y) = \dfrac{c}{x}$ with period $4$. Very similar.
\item $F(x,y) = \dfrac{1}{xy}$ with period $3$.
\item $F(x,y) = \dfrac{y}{x}$ with period $6$.
\item $F(x,y) = x$, with period $2$. 
\end{itemize}

@@

We can probably employ similar strategies for rational functions.
\section{``Raw'' Idea}

>>>>

Suppose there was a function $f$ of one variable such that $F(x,y) = \dfrac{f(y)}{x}$ had periodicity 2, then $a_0 = a_2 = x$, $a_1 = y$ and
\[ a_2 = x = \dfrac{f(y)}{x} \]
And so $f(y) = x^2$, for every $y$, so $f$ is not a function of one variable.

@@

Examples of functions and their periods:
\begin{itemize}
\item $F(x, y) = \dfrac{1}{x}$ with period $4$ as shown above.
\item $F(x,y) = \dfrac{c}{x}$ with period $4$. Very similar.
\item $F(x,y) = \dfrac{1}{xy}$ with period $3$.
\item $F(x,y) = \dfrac{y}{x}$ with period $6$.
\item $F(x,y) = \dfrac{cy}{x}$ with period $6$. 
\item $F(x,y) = x$, with period $2$. 
\end{itemize}

@@

We can probably employ similar strategies for rational functions.

\section{Something}

I can show that if $F(x,y) = \dfrac{f(y)}{x}$, where $f$ is degree greater than or equal to 2, then it will not be periodic.

Suppose $f$ is degree $n$, then by looking at the highest degree terms we can let $y_i$ be the highest degree on the $y$ term of the expression. Then
\[ y_{i+1} = ny_{i} - y_{i-1} \]
\[ y_0 = 0, y_1 = 1 \]
So we can show by induction that $y_{i+1} > y_i$, and so the degree of $y$ never goes to 0 (in order for the formula to evaluate to $x$. Its clear for $i = 0$. Suppose it works for some $i$, then 
\[ y_{i+2} = y_{i+1} + (y_{i+1} - y_{i}) \]
\[ y_{i+2} > y_{i+1} \]
following the inductive hypothesis.

Likewise, if $f(y) = \dfrac{P(y)}{Q(y)}$, where $P$ and $Q$ are polynomials of degree $p$ and $q$, then
\[ y_{i+1} = (p-q)y_{i} - y_{i-1} \]
So if $p - q \geq 2$, then the function is not periodic.

I also came up with a closed form formula for the highest degree of $y$ at the $i$th iteration by a generating function. 
So first, 
\[ Y(x) = \sum y_ix^i = \dfrac{x}{x^2 - nx + 1} \]
And so we can let $a = \dfrac{n + \sqrt{n^2 - 4}}{2}, b = \dfrac{n - \sqrt{n^2-4}}{2}$, and so 
\[ y_i = \dfrac{a}{a^2 - 1}(a^i - b^i) \]

Also, one can ask which periodicty can a function have. Suppose $F(x,y) = f(y)/x$ where $f$ is a polynomial of degree $1$, since degree $0$ we already solved it. Then if we look at the patterns of highest degree terms in $y$, and we have 
\[ y_{i+1} = y_i - y_{i-1} \]
We know that if the function is periodic, then at some $p$, $y_p = 0$ and $y_{p+1} = 1$. This means that $y_{p-1} = -1, y_{p-2} = -1$, then again, $y_{p-3} = 0$. Therefore, we know that in the least, $p$ must be a multiple of $3$. In addition, since we have a change in parity, then we know that $p$ must be a multiple of $6$. In fact, we get $f(y) = y$ gets a periodicity of $6$.

Similarly, if $f(y)$ has highest degree of $y$ with $-1$, then if the function is periodic with periodicity $p$, then $y_p = 0$ and $y_{p+1} = 1$, and so $y_{p-1} = -1$, $y_{p-2} = 1$ and $y_{p-3} = 0$. And so the periodicity must be a multiple of $3$.

\section{``Raw'' Idea}
